### 2021-03-06 chapters/sections/1-context.tex
überarbeitung

Since invention of the World Wide Web, the number of its users increased rapidly. Today, more than 90 percent of the german population uses the internet \cite{Onlinestudie}. Enterprises and organizations recognised this potential to provide their services to a large number of people as Service Providers (SP). A frequently used method of SPs is "online self-service" (OSS). Specialized software tools enable customers to use services through the internet without direct human interaction.

Many tools require identification of a user in case interactions or initiated business processes have to be associated with him. A system for management of identities of users therefore is part of numerous system architectures. Service Providers usually enable users to manage their partial identities using online self-service through user profiles. Today, customers of SPs are associated with multiple partial identities and user profiles - one for each SP.

The result is a variety of problems in management of digital identities \cite{IdentityCrisis}. Here are some examples:
\begin{itemize}
    \item Multiple SPs collect, use and share parts of identities which leads to distributed and fragmented identities.
    \item As a result of the distributed identities, risk of data breaches increases.
    \item Managing multiple user profiles is very inconvenient for users.
    \item It is often unclear who stores which data and for what purpose.
    \item SPs have an incentive to collect and share more user data than they need.
\end{itemize}

An improved model for identity management is necessary. In recent years, multiple new models have been invented. One model called Social Login became popular. It enables customers to create User Profiles for SPs through existing User Profiles of Social Networks like Facebook. This improves identity management by enabling SPs to rely on the external profile for unique identification, authorisation and personal attributes. Users can therefore manage identity related information through their Social Profile which is accessed by the SP.

This approach, however, still leaves the requirement of User Profiles for each SP because, depending on the provided service, additional attributes which are not part of the Social Profile, have to be manageable by the user. SPs also might require additional online self-service tools the social network does not provide.

The solution to this problem is an Identity Management Provider (IMP) which manages the whole online identity of a person and can replace the individual partial identities and user profiles. The system itself is provided and hosted by a SP for consumption by other SPs. To replace existing User Profiles, the IMP can not only rely on systems for registration, authentication, authorisation and provisioning. Additional requirements of SPs like communication, data wallet and management of additional attributes have to be satisfied.

In order for an SP to use identities and services of an IMP, an integration into the system architecture of the SP has to take place. Due to the complex nature of system architectures and identity management, this is a difficult task.

A currently relevant example for usage of online self-service with particular interesting requirements for identities is the German "Online Access Law" (Online Zugangs Gesetz: OZG). It requires all administrative services of the german federal republic, each member state and commune to be digitally available through interoperable user profiles. The current plan is to make the profiles only available for usage in context of the OZG. From a user perspective, this would be yet another partial identity and user profile to manage.

>>>>

Since invention of the World Wide Web, the number of its users increased rapidly. Today, more than 90 percent of the German population uses the internet \cite{Onlinestudie}. Enterprises and organizations recognised this potential to provide their services to a large number of people as Service Providers (SP). A frequently used method of SPs is "online self-service" (OSS). Specialized software tools enable customers to use services through the internet without direct human interaction.

Many tools require identification of users in case interactions or initiated business processes have to be associated with them. A system for management of identities of users therefore is part of numerous system architectures. Service Providers often enable users to manage their partial identities using online self-service through user profiles. Today, customers of SPs are associated with multiple partial identities and user profiles - one for each SP.

The result is a variety of problems in management of digital identities \cite{IdentityCrisis}. Here are some examples:
\begin{itemize}
    \item Multiple SPs collect, use and share parts of identities which leads to distributed and fragmented identities.
    \item As a result of the distributed identities, risk of data breaches increases.
    \item Managing multiple user profiles is very inconvenient. Users have to login to various websites to change attributes of each user profile.
    \item It is often unclear which SP stores which data and for what purpose.
    \item SPs have an incentive to collect and share more user data than they need.
\end{itemize}

An improved identity management is necessary. In recent years, multiple new models have been invented. One model called Social Login became popular. It enables customers to create user profiles for SPs through existing user profiles of Social Networks like Facebook. This improves identity management by enabling SPs to rely on the external profile for unique identification, authorisation and personal attributes. Users can therefore manage identity related information through their Social Profile which is accessed by the SP.

This approach, however, still leaves the requirement of user profiles for each SP because, depending on the provided service, additional attributes which are not part of the Social Profile, have to be manageable by the user.

A possible solution to this problem is Identity Management Provisioning (IMP) which enables users to manage their whole online identity instead of partial identities and user profiles. Trough IMP relationships, users can share personal information of their IMP identity with service providers and exchange data in form of messages. Users remain in control of their identity. They share personal information when needed, retract access for service providers when no longer necessary, securely exchange messages, all trough the same service.

An Identity Management Host (IMH) hosts an Identity Management Provisioning System (IMPS) to make IMP accessible to service providers. In order for a SP to use IMP for identity management, an integration of IMPS into its existing system architecture is necessary. Due to the complex nature of system architectures and identity management, this is a difficult task.

A currently relevant example for usage of online self-service with particular interesting requirements for identity management is the German Online Access Law (Online Zugangs Gesetz - OZG). It requires all administrative services of the German federal republic, each member state and commune to be digitally available through interoperable user profiles. The current plan is to make user profiles only available for usage in context of the OZG. From a user perspective, this would be yet another partial identity and user profile to manage.

### 2021-03-06 chapters/sections/11-technological_integration.tex
überarbeitung

To present an onboarding relationship template to the user, the web server first has to retrieve it from the messaging system. For this purpose, an "Onboarding Template Module" is added to the messaging system. Using the API of a messaging adapter provided in addition to the module, the web server can request the creation of new onboarding templates. Optionally, the web server can pass additional attributes which should be stored in the metadata section of the template. In this case, the web server will want to notify the user about the successful creation of his user profile. The web server therefore adds the users current session ID as metadata to the template.

The messaging adapter is aware of the existence of two channels: the "Onboarding Template Request" channel, where it publishes its requests and the "Onboarding Template Response" channel, where it listens for responses. In order for the adapter to correlate request and reply messages, it creates a message which besides the Template Metadata contains a request ID. This ID is expected to be included in the response message. The message ID is a unique identifier and is part of every message. The module consumes messages from the "Onboarding Template Request" channel. On each channel connected to a messaging adapter, message translators are attached. These translators have the purpose of providing a canonical data model within the module by translating messages on the data representation, data types and data structures layer. It might happen, that in the future one of the adapters switches from using for example an XML instead of a JSON data representation. Using message translators on each side, enables to adapt to these changes by only appropriately configuring the translator while leaving the rest of the module untouched.

@@

The module processes an established relationship by initiating the creation of a new user profile. First, a message filter lets trough only messages of the onboarding type. Afterwards, each message is stored in a "Relationship Storage" database and published on the "Create User Profile Request" channel. However, before publishing the message, a message translator maps attributes form the IMP data model to the OZG data model.

@@

Trough the "New Relationship Mapping" channel, the "Relationship Mapping" module receives the message sent by the previously described module. Purpose of the "Relationship Mapping" module is to map relationships to corresponding entities in the existing system architecture. In this case, relationships are mapped to user profiles and vice versa using relationship IDs and user profile IDs. The message consumed trough the "New Relationship Mapping" channel is stored in a database for enabling future translation between the contained relationship ID and user profile ID. Using the "Relationship Mapping Request" channel, modules can request the "Relationship Mapping" module to add a user profile ID for a given relationship ID and vice versa.

@@

Incoming messages on the "Message Inbound" channel are consumed by the module and translated by a message translator to a canonical data model. In order for the module to identify a response after publishing the message on the "Relationship Mapping Request" channel, a content enricher adds a unique request ID to the message and stores the ID in a database. Only messages on the "Relationship Mapping Response" channel containing a request ID stored in the database are let trough by the request ID message filter. As the following systems will process the inserted user profile ID instead of the relationship ID, a content filter removes relationship IDs from the message before publishing it on the "Message Inbound Mapping" channel.

Outbound messages on the "Message Outbound Mapping" channel are processed similar to inbound messages. The difference being, that the user profile ID contained in the message is replaced by the corresponding relationship ID.

@@

In this case, an "Attribute Change Message Module" is included, which processes messages related to the attribute change process. Messages exchanged trough IMP relationships can have arbitrary content, however, the messages are assumed to contain a message type attribute and a message content attribute. Based on the message type attribute, a message filter can let messages trough which contain a message type attribute correlating to an attribute change process. Based on the type of a message, a content based message router then copies each message to a datatype channel. In this case, messages of type attribute change request and attribute change established are distinguished. Corresponding to that, two data type channels exist. As based on the channel, a message is placed, its message type can be inferred, content filters remove the message type attributes.

@@

To process incoming attribute change requests for profiles, a "User Profile Attribute Change Request" module is included which consumes messages from the "Attribute Change Request" channel. Messages arriving on the "Attribute Change Request" channel were mapped form a relationship ID to a user profile ID in an earlier integration step. Depending on the use case of the service provider, not all user profiles should be processed in the same way. Therefore, message filters are inserted, which are able to distinguish possible "types" of user profiles based on the user profile ID. It would also be possible to include an additional module, which is able to request user profile type information from the existing systems, separating messages on the "Attribute Change Request" channel appropriately. For example, user profiles for businesses could exist, where processing of attribute change requests requires a more through and manual verification by employees.

@@

If the user did not cancel the attribute change request before the OZG system accepted, the attribute change is established. The connector is notified trough a message, which eventually by processed by the "User Profile Attribute Change Message" module described earlier and published on the "Attribute Change Established" channel.

@@

In order for the user to receive messages sent to his user profile trough the IMP client, mail messages are processed by the integration architecture. Similar to the "Profile Attribute Change Message" module, a "Mail Message" module is included, which maps messages on the "Mail Outbound" channel to the "Message Outbound Mapping" channel.

@@

A messaging adapter attached to the service facade is notified each time, a user profile connected to an IMP identity receives a mail and publishes a message on the "Profile Mail Outbound" channel. The message contains the ID of the profile which received the message and the content of the mail.

@@

If the user creates a user profile using his IMP identity, it is possible to replace authentication using a password with authentication trough the IMP client. It is also possible to use the IMP client for two factor authentication.

@@

When the user visits the web page of the administration portal to apply for an administrative service, he will have to login. To utilize the IMP client for authentication, the web server can use a channel adapter to publish a authentication request concerning a user profile ID and a certain session ID. The adapter publishes a message on the "Authenticate" channel which contains the profile ID and a request content. Next to the session ID, the request content may also include information for presentation to the user like the current time or the IP address of the browser. Similar to previous modules, a message translator is inserted to translate the message to a canonical data model and publish it on the "Authentication Request Message".

>>>>

To present an onboarding relationship template to the user, the web server first has to retrieve it from the messaging system. For this purpose, an "Onboarding Template Module" is added to the messaging system. Using the API of a messaging adapter provided in addition to the module, the web server can request the creation of new onboarding templates. Optionally, the web server can pass additional attributes which should be stored in the metadata section of the template. In this case, the web server will want to notify the user about the successful creation of his user profile. The web server therefore adds the user's current session ID as metadata to the template.

The messaging adapter is aware of the existence of two channels: the "Onboarding Template Request" channel, where it publishes its requests and the "Onboarding Template Response" channel, where it listens for responses. In order for the adapter to correlate request and reply messages, it creates a message which besides the Template Metadata contains a request ID. This ID is expected to be included in the response message. The message ID is a unique identifier and is part of every message. The module consumes messages from the "Onboarding Template Request" channel. On each channel connected to a messaging adapter, message translators are attached. These translators have the purpose of providing a canonical data model within the module by translating messages on the data representation, data types and data structures layer. It might happen, that in the future one of the adapters switches from using for example an XML instead of a JSON data representation. Using message translators on each side enables to adapt to these changes by only appropriately configuring the translator while leaving the rest of the module untouched.

@@

The module processes an established relationship by initiating the creation of a new user profile. First, a message filter lets through only messages of type "onboarding". Afterwards, each message is stored in a "Relationship Storage" database and published on the "Create User Profile Request" channel. However, before publishing the message, a message translator maps attributes from the IMP data model to the OZG data model.

@@

Through the "New Relationship Mapping" channel, the "Relationship Mapping" module receives the message sent by the previously described module. Purpose of the "Relationship Mapping" module is to map relationships to corresponding entities in the existing system architecture. In this case, relationships are mapped to user profiles and vice versa using relationship IDs and user profile IDs. The message consumed through the "New Relationship Mapping" channel is stored in a database for enabling future translation between the contained relationship ID and user profile ID. Using the "Relationship Mapping Request" channel, modules can request the "Relationship Mapping" module to add a user profile ID for a given relationship ID and vice versa.

@@

Incoming messages on the "Message Inbound" channel are consumed by the module and translated by a message translator to a canonical data model. In order for the module to identify a response after publishing the message on the "Relationship Mapping Request" channel, a content enricher adds a unique request ID to the message and stores the ID in a database. Only messages on the "Relationship Mapping Response" channel containing a request ID stored in the database are let through by the request ID message filter. As the following systems will process the inserted user profile ID instead of the relationship ID, a content filter removes relationship IDs from the message before publishing it on the "Message Inbound Mapping" channel.

Outbound messages on the "Message Outbound Mapping" channel are processed similarly to inbound messages. The difference is that the user profile ID contained in the message is replaced by the corresponding relationship ID.

@@

In this case, an "Attribute Change Message Module" is included, which processes messages related to the attribute change process. Messages exchanged through IMP relationships can have arbitrary content. However, the messages are assumed to contain a message type attribute and a message content attribute. Based on the message type attribute, a message filter can let messages through which contain a message type attribute correlating to an attribute change process. Based on the type of a message, a content based message router then copies each message to a datatype channel. In this example, messages of type attribute change request and attribute change established are distinguished. Corresponding to that, two data type channels exist. As based on the channel, a message is placed, its message type can be inferred, content filters remove the message type attributes.

@@

To process incoming attribute change requests for profiles, a "User Profile Attribute Change Request" module is included which consumes messages from the "Attribute Change Request" channel. Messages arriving on the "Attribute Change Request" channel were mapped from a relationship ID to a user profile ID in an earlier integration step. Depending on the use case of the service provider, not all user profiles should be processed in the same way. Therefore, message filters are inserted, which are able to distinguish possible "types" of user profiles based on the user profile ID. It would also be possible to include an additional module, which is able to request user profile type information from the existing systems, separating messages on the "Attribute Change Request" channel appropriately. For example, user profiles for businesses could exist, where processing of attribute change requests requires a more through and manual verification by employees.

@@

If the user did not cancel the attribute change request before the OZG system accepted, the attribute change is established. The connector is notified through a message, which eventually is processed by the "User Profile Attribute Change Message" module described earlier and published on the "Attribute Change Established" channel.

@@

In order for the user to receive messages sent to his user profile through the IMP client, mail messages are processed by the integration architecture. Similar to the "Profile Attribute Change Message" module, a "Mail Message" module is included, which maps messages on the "Mail Outbound" channel to the "Message Outbound Mapping" channel.

@@

A messaging adapter attached to the service facade is notified each time a user profile connected to an IMP identity receives a mail and publishes a message on the "Profile Mail Outbound" channel. The message contains the ID of the profile which received the message and the content of the mail.

@@

If the user creates a user profile using his IMP identity, it is possible to replace password authentication with authentication through the IMP client. It is also possible to use the IMP client for two factor authentication.

@@

When the user visits the web page of the administration portal to apply for an administrative service, he will have to login. To utilize the IMP client for authentication, the web server can use a channel adapter to publish an authentication request concerning a user profile ID and a certain session ID. The adapter publishes a message on the "Authenticate" channel which contains the profile ID and a request content. Next to the session ID, the request content may also include information for presentation to the user like the current time or the IP address of the browser. Similar to previous modules, a message translator is inserted to translate the message to a canonical data model and publish it on the "Authentication Request Message".

### 2021-03-06 chapters/sections/13-imp_solution_integration.tex
überarbeitung

In contrast to relationship templates in the previous IMP solution, these wont have to be dynamically created but only once. The reason is, that in the previous IMP solution templates had to contain personalized information like the session ID. This is not necessary in this case.

Once the templates are created, the corresponding template IDs can be rendered as QR code on the web page of the institution. If the content of a template has to be changed, an employee can be manually create a new template and switch the template IDs on the web page.

Visiting the web page of the institution, the user can scan the QR code and is presented with the relationship template. The user can read trough the included description of the corresponding administrative service to determine if it is the correct one. Based on the requested attributes and the information of how they are processed, the user can make an informed decision whether to apply for the administrative service.

@@

When accepting the relationship and therefore the application request, and the user did not retract his request in the meantime, the relationship is established and the institution gets notified of a new application.

After the administrative service is finished, the relationship can be terminated and the institution loose access to all shared personal information.

@@

This IMP solution utilizes the same message types as the previous IMP solution however with a different purpose.

@@

Trough the IMP client users can request attribute changes for established relationships. The purpose of these messages is to enable uses to change the content of an application even after it was submitted.

@@

As in the previous IMP solution, users should be able to receive mails from every institution trough the IMP client. Each mail is associated to a relationship, enabling the user to easily correlate mails to active applications.

@@

The IMP solution enables users to share personal information directly with responsible institution, restricting access to for example administration portals. It also enables direct and secure exchange of messages.
As a result of the direct mapping of relationships to applications, users can request the modification of application content even after submission.

>>>>

In contrast to relationship templates in the previous IMP solution, these wont have to be dynamically created but only once. The reason is that in the previous IMP solution templates had to contain personalized information like the session ID. This is not necessary in this case.

Once the templates are created, the corresponding template IDs can be rendered as QR code on the web page of the institution. If the content of a template has to be changed, an employee can manually create a new template and switch the template IDs on the web page.

Visiting the web page of the institution, the user can scan the QR code and is presented with the relationship template. The user can read through the included description of the corresponding administrative service to determine if it is the correct one. Based on the requested attributes and the information on how they are processed, the user can make an informed decision whether to apply for the administrative service.

@@

When accepting the relationship and therefore the application request, and the user did not retract his request in the meantime, the relationship is established and the institution gets notified about a new application.

After the administrative service is finished, the relationship can be terminated and the institution looses access to all shared personal information.

@@

This IMP solution utilizes the same message types as the previous IMP solution, however with a different purpose.

@@

Through the IMP, client users can request attribute changes for established relationships. The purpose of these messages is to enable users to change the content of an application even after it was submitted.

@@

As in the previous IMP solution, users should be able to receive mails from every institution through the IMP client. Each mail is associated with a relationship, enabling the user to easily correlate mails to active applications.

@@

The IMP solution enables users to share personal information directly with responsible institutions, restricting access to for example administration portals. It also enables direct and secure exchange of messages.
As a result of the direct mapping of relationships to applications, users can request the modification of application content even after submission.

### 2021-03-06 chapters/sections/14-technological_integration.tex
überarbeitung

Institutions are able to process a predetermined set of administrative services. For each service, the institution can once create a relationship template and store them as QR code on their web page.

@@

The layout of the messaging system is identical to the one presented in section 4.2. Exactly the same type of messages are exchanged with however, a slightly different purpose and content. Therefore, in this and the following sections, only important differences between the operation of the messaging system will be described.

On the "Relationship to Identity Integration Layer", modules exist for establishing the different types of application relationships. For simplification, modules only for an "Application A" are shown. Besides the content and purpose of relationships in this case are different to the relationships of chapter 4, they are processed exactly the same: Each type of relationship request is processed by an individual application request module and each type of established relationship is processed by an individual application relationship module. As the purpose of relationships in this case are applications for administrative services, each module has to be configured appropriately. For an established relationship, instead of issuing the existing system architecture to create a user profile, the application relationship module issues the existing system architecture to process an application based on the attributes shared as part of the relationship. And instead of a user profile ID, the existing system architecture responds with an application process ID. This application process ID is used in the same way as the user profile ID in section 4.2. Trough the "New Relationship Mapping" channel, it is sent to the relationship mapping module, which maps each relationship to an application process and vice versa.

On the "Communication Integration Layer", the same types of messages are exchanged. However instead of user profile IDs, relationship IDs are mapped to application process IDs. In chapter 4, each module connected to the messaging adapter of the institution system was designed to be able to distinguish between different types of user profile IDs in order to for example distinguish a enterprise profile from the profile of a private person. In this case, modules can distinguish between different application types based on the application process ID.

@@

Relationship templates are created manually and stored as a QR-code image on the web page of the institution. Fore each application type, a separate application template module can be included which stores the configuration for the template type and issues the connector to create new relationship templates. Instead of the web page, the institution system is connected, as application templates are not dynamically created for user requests but once by an employee of the institution. Each time, the content of an application template has to be changed, the institution creates a new one trough this module.

@@

For each type of application relationship, a different application request module is used to separate the processing of requests for different application types into application request processing modules. Depending on the type of application, different requirements regarding the acceptance of application requests exist. A request processing module might for example check if the home address of the user lies in the are of responsibility of the institution. Only after the relationship and therefore the application request is accepted will the application be processed.

@@

For each type of application relationship a different application relationship module is used for processing established relationships. The institution system is notified of an accepted application trough the "Process Application Request" channel. As part of the existing system architecture, institution systems received applications trough the data exchange platform as an application document. In order to adhere to this standard, the message translator connected to the "Process Application Request" channel translates the relationship content into the application format of documents delivered trough the data exchange platform. As explained in section 2.3.2, form servers create the application documents based on a standard defined by FIM.

Institution systems are assumed to uniquely reference active applications with process IDs. After starting the application process, the institution responds trough the messaging adapter with the process ID on the "Application Process Reference Response" channel. This process ID is then published on the "New Relationship Mapping" channel. The result is, that in contrast to the messaging system of chapter 4, relationships are not mapped to user profiles but to application processes.

@@

In the same way as in chapter 4, attribute change message types are published on individual channels, this time however, instead of a user profile ID, they contain a process ID. As described on the example of different user profile ID types in chapter 4, different process ID types may exist in this case too. For example, an attribute change request for an application of type "A" might be processed differently to an attribute change request for an application of type "B". To process different types of attribute change requests separately, multiple application attribute change request modules can be included, each of them filtering for different types of process IDs. If it is possible to retrieve the application type from the process ID, processing of each application type could be separated into multiple application change request modules. In case, application types are not retrievable from process IDs, an additional module can be included, which requests the application type from the institution system based on the process ID. An alternative would be for the application relationship module to not only publish the process ID corresponding to a relationship ID on the "New Relationship Mapping" channel but also the application type.

Only after the attribute change request is accepted, will the institution system be notified of the attribute change.

@@

The institution system receives the notification of a modification of the application trough the messaging adapter. Depending on the capabilities of the institution system, this could result in either an automated update of the referenced application process or an automated E-Mail to the assigned employee.

@@

Similar to the attribute change messages, mail messages can be fitted to the process based scenario, enabling institutions to send mails to a process ID which the messaging system translates to a relationship ID in order for the correct IMP identity to receive it. In addition to the one directional message transfer from the institution to the IMP identity, the application mail module and the mail message module could be expanded to be able to receive mails from an IMP identity trough an established relationship. The relationship ID of the message would be mapped to the corresponding process ID and submitted it to the institution system. The institution system could then send the mail to the responsible employee.

@@

It was also able to integrate a user oriented (persistent) relationship scenario as well as a process oriented (temporary) relationship scenario. This demonstrates, that the presented integration architecture can be used for integrating a variety of IMP solutions.

>>>>

Institutions are able to process a predetermined set of administrative services. For each service, the institution can create a relationship template once and store it as QR code on their web page.

@@

The layout of the messaging system is identical to the one presented in section 4.2. Exactly the same type of messages are exchanged with however a slightly different purpose and content. Therefore, in this and the following sections, only important differences between the operation of the messaging system will be described.

On the "Relationship to Identity Integration Layer", modules exist for establishing the different types of application relationships. For simplification, modules only for an "Application A" are shown. Besides the content and purpose of relationships in this case are different to the relationships of chapter 4, they are processed exactly the same: each type of relationship request is processed by an individual application request module and each type of established relationship is processed by an individual application relationship module. As the purpose of relationships in this case are applications for administrative services, each module has to be configured appropriately. For an established relationship, instead of issuing the existing system architecture to create a user profile, the application relationship module issues the existing system architecture to process an application based on the attributes shared as part of the relationship. And instead of a user profile ID, the existing system architecture responds with an application process ID. This application process ID is used in the same way as the user profile ID in section 4.2. Through the "New Relationship Mapping" channel, it is sent to the relationship mapping module, which maps each relationship to an application process and vice versa.

On the "Communication Integration Layer", the same types of messages are exchanged. However instead of user profile IDs, relationship IDs are mapped to application process IDs. In chapter 4, each module connected to the messaging adapter of the institution system was designed to be able to distinguish between different types of user profile IDs in order to for example distinguish an enterprise profile from the profile of a private person. In this case, modules can distinguish between different application types based on the application process ID.

@@

Relationship templates are created manually and stored as a QR-code image on the web page of the institution. For each application type, a separate application template module can be included which stores the configuration for the template type and issues the connector to create new relationship templates. Instead of the web page, the institution system is connected, as application templates are not dynamically created for user requests but once by an employee of the institution. Each time the content of an application template has to be changed, the institution creates a new one through this module.

@@

For each type of application relationship, a different application request module is used to separate the processing of requests for different application types into application request processing modules. Depending on the type of application, different requirements regarding the acceptance of application requests exist. A request processing module might for example check if the home address of the user is within the are of responsibility of the institution. Only after the relationship and therefore the application request is accepted the application will be processed.

@@

For each type of application relationship a different application relationship module is used for processing established relationships. The institution system is notified of an accepted application through the "Process Application Request" channel. As part of the existing system architecture, institution systems received applications through the data exchange platform as an application document. In order to adhere to this standard, the message translator connected to the "Process Application Request" channel translates the relationship content into the application format of documents delivered through the data exchange platform. As explained in section 2.3.2, form servers create the application documents based on a standard defined by FIM.

Institution systems are assumed to uniquely reference active applications with process IDs. After starting the application process, the institution responds through the messaging adapter with the process ID on the "Application Process Reference Response" channel. This process ID is then published on the "New Relationship Mapping" channel. The result is, that in contrast to the messaging system of chapter 4, relationships are not mapped to user profiles but to application processes.

@@

In the same way as in chapter 4, attribute change message types are published on individual channels, this time however, instead of a user profile ID, they contain a process ID. As described on the example of different user profile ID types in chapter 4, different process ID types may exist in this case too. For example, an attribute change request for an application of type "A" might be processed differently to an attribute change request for an application of type "B". To process different types of attribute change requests separately, multiple application attribute change request modules can be included, each of them filtering for different types of process IDs. If it is possible to retrieve the application type from the process ID, processing of each application type could be separated into multiple application change request modules. In case application types are not retrievable from process IDs, an additional module can be included, which requests the application type from the institution system based on the process ID. An alternative would be for the application relationship module to not only publish the process ID corresponding to a relationship ID on the "New Relationship Mapping" channel but also the application type.

Only after the attribute change request is accepted, the institution system will be notified about the attribute change.

@@

The institution system receives the notification from the application through the messaging adapter. Depending on the capabilities of the institution system, this could result in either an automated update of the referenced application process or an automated e-mail to the assigned employee.

@@

Similar to the attribute change messages, mail messages can be fitted to the process based scenario, enabling institutions to send mails to a process ID which the messaging system translates to a relationship ID in order for the correct IMP identity to receive it. In addition to the one directional message transfer from the institution to the IMP identity, the application mail module and the mail message module could be expanded to be able to receive mails from an IMP identity through an established relationship. The relationship ID of the message would be mapped to the corresponding process ID and submitted to the institution system. The institution system could then send the mail to the responsible employee.

@@

It was also able to integrate a user oriented (persistent) relationship scenario as well as a process oriented (temporary) relationship scenario. This demonstrates that the presented integration architecture can be used for integrating a variety of IMP solutions.

### 2021-03-06 chapters/sections/5-imp.tex
überarbeitung

\begin{itemize}
    \item IMP hosts systems for identity management and provides access to users with an application and to service providers with integration tools
    \item Users can create an identity and manage attributes (key-value pairs)
    \item Users and service providers can establish relationships and use them to securely share personal information and exchange structured data
\end{itemize}

\begin{itemize}
    \item Ease of Use: Users only need to manage one online identity which can be accessed by multiple service providers
    \item Data protection: Users can control which service providers get access to which attributes
    \item Bidirectional Communication: Service providers and users can directly communicate with each other by exchanging structured data their tools have a common understanding of.
    \item User Interaction: The user application can display various user interfaces based on the content of exchanged structured data.
    \item Data Abuse Protection: Service providers are only supplied with necessary personal information which minimizes the possibilities of abuse by the service provider or in case of a data breach.
    \item Communication Security: Attributes communicated structured data are securely exchanged. User and service provider can therefore also directly exchange sensible information. This is not possible for example through e-mail
\end{itemize}

The purpose of the following sections is not to give a detailed explanation of the presented IMP system but to deliver a brought understanding of the concepts, features and components involved.

\subsection{Features}

\paragraph{Provisioning} The IMP system is hosted on the system architecture of the identity management provider and provides identity management as a service for users and SPs. Users and service providers can manage identities by accessing features of the IMP system. The IMP system stores all personal information and data necessary for the operation of identity management. The user only manages one identity which can be accessed by multiple service providers.

\paragraph{Applications} The IMP system provides users and service providers with software tools for accessing its features. The user is provided with a user friendly application while the enterprise is provided with tools for integration.

@@

\subsection{Evaluation of Features}

This section evaluates the advantages and disadvantages of the described features of the IMP system based on the requirements of users and service providers.

\subsubsection{User}
The IMP system enables users to manage one identity and share attributes with multiple service providers. This replaces the need of a user profile for all service providers which take advantage of the IMP service. The user wont have to access multiple websites for managing user profiles but update his identity only through the IMP service.
As attributes can only be shared with service providers as part of a relationship, the user has control of his personal information. Before accepting the relationship, he can see information about the identity of the service provider, like his name, address, phone number and more. The user is also provided with information on why the relationship and request attributes are necessary which helps the user in deciding if he agrees with the necessity of sharing the information. As it is possible to establish multiple relationships with a service provider, each with a different specified reason and list of shared attributes, it is possible for user and service provider to establish a relationship for each business process which requires access to personal information. The information therefore can not only be shared with a single service provider but also for an individual process. User and service provider can interact through relationships. Therefore messages can be exchanged and associated to a certain business process.

\subsubsection{Service Provider}

The IMP system enables the service provider to access personal information of users without the need of hosting a system for identity management. Depending on the price of the IMP solution, this can reduce the cost for operating and maintaining the system architecture. The service provider has to hire less staff, does not have to buy expensive hardware and can more easily scale the performance of his system. As previously described, the user experience can also increase and lead to more customers. Relationships enable service providers to request users to share attributes while specifying the legal reasons for processing the information and to directly interact with them. The IMP system takes over the responsibility for securing personal information of users and interactions with the service provider. As the relationship cannot be deleted without approval of both identities, it can be used as proof of for example contracts or order placements. The IMP system provides tools for integration which minimizes the effort for system providers to start using the IMP system.

As the IMP system focuses on improved data protection for users, the service provider might lose access to valuable but not necessary personal information. The determination of a reason for processing personal information also requires increased system complexity for making sure, that personal information is only used for the specified reason. The integration of a new solution for identity management is an expensive and time consuming task and might in short term increase instead of reduce cost. Using an external service for identity management makes the service provider dependent on the identity management provider. This can be a problem, if in the future, the service costs increase or the IMP solution stops being supported.

\subsection{Components}

@@

\begin{figure}[h]
\caption{IMP System Operation Layers}
    \centering
    \includegraphics[scale=0.3]{Diagrams/IMP Layer Diagram.png}
\end{figure}

>>>>

The purpose of the following sections is not to give a detailed explanation but to deliver a brought understanding of what identity management provisioning is and how identity management provisioning systems operate.

Identity management provisioning (IMP) is a service which enables users to manage a digital identity and use it to share personal information and interact with service providers.

Any user can create an IMP identity and similar to user profiles, add any personal information as attributes. In contrast to existing user profile based identity management systems, however, only one IMP identity and only one identity management provisioning system (IMPS) are necessary. The IMPS is hosted by an identity management provisioning host (IMPH) to enable every service provider to access its service.

Service providers can use the service for identity management by requesting access to personal information of IMP identities. The user has the option of accepting or rejecting to share personal information. As part of the service, users and service provider are also enabled to exchange messages to interact with each other.

\subsection{IMP Example}

This section describes what an identity management provisioning service can look like and is based on the identity management provisioning system called IDAS.

\paragraph{Applications} Users and service providers can access  software tools for accessing its features. The user is provided with a user friendly application while the enterprise is provided with tools for integration.

@@

\subsection{IMPS Example}

@@

\begin{figure}[h]
\caption{IMP System Operation Layers}
    \centering
    \includegraphics[scale=0.6]{Diagrams/IMP Layer Diagram.pdf}
\end{figure}
